---
title: "Harmonized Drug Dataset Quality Control"
author: "James Bannon, Michael Tran"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE,
  fig.width = 7,
  fig.height = 5
)
knitr::opts_knit$set(root.dir = normalizePath(".."))
options(digits = 3)
```

# Overview

This report is a quality control walkthrough for the Harmonized Drug Dataset
stored as a `MultiAssayExperiment` (MAE). The dataset harmonizes multiple drug-related
assays and features across a shared compound universe keyed by `HDD.Compound.ID` so
downstream analyses can operate on a consistent sample space. `Pubchem.CID` is retained
when available, but it is not required for a compound to be included.

## Data Sources

The MAE is constructed in `workflow/scripts/construct_MAE.R`. It pulls processed experiments
from `data/procdata/experiments/` and a `colData` table from `data/procdata/colData.csv`.
Key sources (from `config/pipeline.yaml` and the workflow scripts) include:

- **DeepChem datasets** (Tox21, ToxCast, SIDER, ClinTox): downloaded from DeepChem S3, then re-shaped to `HDD.Compound.ID`-by-assay matrices.
- **JUMP-CP** membership and source identifiers from the curated cpg0016 MAE.
- **OASIS** membership and source identifiers from the curated cpg0037 MAE.
- **GEOM** membership and source identifiers from the curated GEOM MAE.
- **LINCS** membership and source identifiers from the curated LINCS MAE.
- **Bioassays**: pulled from AnnotationDB alongside compound metadata, filtered to gold-standard assays.
- **Fingerprints**: RDKit Morgan count fingerprints generated from SMILES.

# Load HDD

```{r load-data}
library(MultiAssayExperiment)
library(SummarizedExperiment)
library(Matrix)
library(ggplot2)
library(dplyr)
library(tidyr)

hdd_mae <- readRDS("data/results/HDD_v2.RDS")
hdd_mae
```

```{r run-artifacts}
format_file_size <- function(bytes) {
  if (is.na(bytes)) {
    return(NA_character_)
  }

  units <- c("B", "KB", "MB", "GB")
  unit_idx <- 1
  value <- bytes
  while (value >= 1024 && unit_idx < length(units)) {
    value <- value / 1024
    unit_idx <- unit_idx + 1
  }

  sprintf("%.1f %s", value, units[[unit_idx]])
}

expected_outputs <- c(
  "data/results/HDD_v2.RDS",
  "data/results/HDD_v2_csv",
  "data/results/HDD_v2_csv.tar.gz"
)

output_summary <- tibble::tibble(
  path = expected_outputs,
  exists = file.exists(expected_outputs),
  size = vapply(
    expected_outputs,
    function(path) {
      if (file.exists(path) && !dir.exists(path)) {
        file.info(path)$size
      } else if (dir.exists(path)) {
        sum(
          file.info(list.files(path, recursive = TRUE, full.names = TRUE))$size,
          na.rm = TRUE
        )
      } else {
        NA_real_
      }
    },
    numeric(1)
  ),
  size_pretty = vapply(size, format_file_size, character(1)),
  modified = as.POSIXct(
    vapply(
      expected_outputs,
      function(path) {
        if (file.exists(path)) {
          file.info(path)$mtime
        } else {
          NA
        }
      },
      as.POSIXct(NA)
    ),
    origin = "1970-01-01"
  )
)

output_summary
```

```{r compound-labels}
compound_labels <- as.data.frame(colData(hdd_mae)) |>
  tibble::rownames_to_column("sample_id") |>
  transmute(
    sample_id = as.character(sample_id),
    compound_label = dplyr::coalesce(
      na_if(trimws(as.character(Molecule.Name)), ""),
      na_if(trimws(as.character(IUPAC.Name)), ""),
      na_if(trimws(as.character(Pubchem.CID)), ""),
      na_if(trimws(as.character(HDD.Compound.ID)), ""),
      sample_id
    )
  )

compound_labels$compound_label <- make.unique(compound_labels$compound_label)
```

# MAE snapshot

```{r mae-snapshot}
experiment_names <- names(experiments(hdd_mae))
experiment_names

has_experiment <- function(name) {
  name %in% experiment_names
}

summarize_experiment <- function(se) {
  assay_obj <- assays(se)[[1]]
  data_class <- class(assay_obj)[1]
  dims <- dim(assay_obj)
  missing_fraction <- if (is.numeric(assay_obj) || is.logical(assay_obj)) {
    mean(is.na(assay_obj))
  } else {
    mean(is.na(as.vector(assay_obj)))
  }

  data.frame(
    assay_name = assayNames(se)[1],
    rows = dims[1],
    cols = dims[2],
    data_class = data_class,
    missing_fraction = missing_fraction,
    stringsAsFactors = FALSE
  )
}

experiment_summary <- do.call(
  rbind,
  lapply(experiments(hdd_mae), summarize_experiment)
)
rownames(experiment_summary) <- experiment_names
experiment_summary
```

Basic sanity checks for sampleMap completeness and the HDD v2 schema.

```{r samplemap-checks}
# Basic sampleMap sanity checks
all_in_coldata <- lapply(experiments(hdd_mae), function(se) {
  all(colnames(se) %in% rownames(colData(hdd_mae)))
})
if (all(unlist(all_in_coldata))) {
  print("All experiments map cleanly to colData.")
} else {
  print("Some experiments contain samples missing from colData:")
  print(all_in_coldata[!unlist(all_in_coldata)])
}

# Any missing samples per experiment (only report issues)
missing_samples <- lapply(experiments(hdd_mae), function(se) {
  setdiff(colnames(se), rownames(colData(hdd_mae)))
})
missing_samples <- missing_samples[lengths(missing_samples) > 0]
if (length(missing_samples) == 0) {
  print("No missing samples detected in any experiment.")
} else {
  print("Missing samples by experiment:")
  print(missing_samples)
}

schema_checks <- tibble::tibble(
  check = c(
    "LINCS membership present",
    "JUMP-CP membership present",
    "OASIS membership present",
    "GEOM membership present",
    "HDD compound ID present"
  ),
  passed = c(
    "In.LINCS" %in% colnames(colData(hdd_mae)),
    "In.JUMP.CP" %in% colnames(colData(hdd_mae)),
    "In.OASIS" %in% colnames(colData(hdd_mae)),
    "In.GEOM" %in% colnames(colData(hdd_mae)),
    "HDD.Compound.ID" %in% colnames(colData(hdd_mae))
  )
)
schema_checks
stopifnot(all(schema_checks$passed))
```

# colData overview

```{r coldata-overview}
cd <- colData(hdd_mae)
cd[1:5, 1:min(10, ncol(cd))]

# Quick missingness scan
missing_by_field <- sapply(cd, function(x) {
  mean(is.na(x))
})
missing_by_field
```

```{r coldata-basic-stats}
# Distinct counts for key identifiers
c(
  n_compounds = nrow(cd),
  unique_hdd_ids = length(unique(cd$HDD.Compound.ID)),
  unique_non_missing_cids = length(unique(na.omit(cd$Pubchem.CID)))
)
```

# Experiment walkthroughs

Below, each experiment in the HDD MAE is reviewed.

## Bioassays

A compound-by-assay matrix of PubChem bioassay outcomes. Rows are AIDs
(assay identifiers) and columns are `HDD.Compound.ID` values. Entries are categorical outcomes,
with missing values representing compounds that were not measured.

```{r bioassays-preview}
se <- experiments(hdd_mae)[["Bioassays"]]
assay_mat <- assay(se)

c(rows = nrow(assay_mat), cols = ncol(assay_mat))
assay_mat[1:min(5, nrow(assay_mat)), 1:min(6, ncol(assay_mat))]
```

```{r bioassays-qc}
# Category counts
assay_mat_chr <- as.matrix(assay_mat)
assay_values <- ifelse(
  is.na(assay_mat_chr),
  "Not Measured",
  as.character(assay_mat_chr)
)
category_counts <- sort(table(assay_values), decreasing = TRUE)
category_counts

inactive_count <- if ("Inactive" %in% names(category_counts)) {
  category_counts[["Inactive"]]
} else {
  0
}
cat("Inactive values in current Bioassays matrix:", inactive_count, "\n")

# Missingness proxy: Not Measured rate per assay
not_measured_rate <- rowMeans(is.na(assay_mat_chr))
summary(not_measured_rate)
```

```{r bioassays-plots, fig.height=6, fig.width=8}
bio_counts <- tibble::tibble(
  status = factor(
    names(category_counts),
    levels = c("Active", "Inactive", "Not Measured")
  ),
  count = as.numeric(category_counts)
)

bio_by_assay <- t(apply(assay_mat_chr, 1, function(values) {
  table(factor(
    ifelse(is.na(values), "Not Measured", as.character(values)),
    levels = c("Active", "Inactive", "Not Measured")
  ))
})) |>
  as.data.frame() |>
  tibble::rownames_to_column("assay") |>
  tidyr::pivot_longer(-assay, names_to = "status", values_to = "count")

bio_by_assay$status <- factor(
  bio_by_assay$status,
  levels = c("Active", "Inactive", "Not Measured")
)

ggplot(bio_counts, aes(x = status, y = count)) +
  geom_col(fill = "#4C78A8") +
  labs(
    title = "Bioassays overall category counts",
    x = "Outcome",
    y = "Count"
  ) +
  theme_minimal()

ggplot(bio_by_assay, aes(x = assay, y = count, fill = status)) +
  geom_col(position = "fill") +
  coord_flip() +
  labs(
    title = "Bioassays outcome composition by assay",
    x = "Assay (AID)",
    y = "Fraction",
    fill = ""
  ) +
  theme_minimal()
```

## Tox21

A compound-by-assay matrix for the Tox21 panel. Rows are Tox21 assays and
columns are `HDD.Compound.ID` values. Values are typically binary (0/1) with missing values for
unmeasured compounds.

```{r tox21-preview}
se <- experiments(hdd_mae)[["Tox21"]]
assay_mat <- assay(se)

c(rows = nrow(assay_mat), cols = ncol(assay_mat))
assay_mat[1:min(5, nrow(assay_mat)), 1:min(6, ncol(assay_mat))]
```

```{r tox21-qc}
# Value distribution
assay_mat_num <- as.matrix(assay_mat)
if ("mol_id" %in% rownames(assay_mat_num)) {
  assay_mat_num <- assay_mat_num[
    rownames(assay_mat_num) != "mol_id",
    ,
    drop = FALSE
  ]
}
table(as.vector(assay_mat_num), useNA = "ifany")

# Missingness per assay
summary(rowMeans(is.na(assay_mat_num)))
```

```{r tox21-plots, fig.height=5, fig.width=8}
tox21_long <- as.data.frame(assay_mat_num) |>
  tibble::rownames_to_column("assay") |>
  tidyr::pivot_longer(-assay, names_to = "compound", values_to = "value")
tox21_long$value_factor <- factor(tox21_long$value, levels = c(0, 1))

tox21_compound_summary <- tox21_long |>
  group_by(compound) |>
  summarize(positive_count = sum(value == 1, na.rm = TRUE), .groups = "drop") |>
  left_join(compound_labels, by = c("compound" = "sample_id")) |>
  arrange(positive_count, compound)

tox21_plot_compounds <- bind_rows(
  tox21_compound_summary |> slice_head(n = 20) |> mutate(group = "Bottom 20"),
  tox21_compound_summary |> slice_tail(n = 20) |> mutate(group = "Top 20")
) |>
  distinct(compound, .keep_all = TRUE) |>
  arrange(positive_count, compound)

tox21_plot <- tox21_long |>
  filter(compound %in% tox21_plot_compounds$compound) |>
  left_join(
    tox21_plot_compounds |>
      select(compound, compound_label),
    by = "compound"
  ) |>
  mutate(
    compound_label = factor(
      compound_label,
      levels = tox21_plot_compounds$compound_label
    )
  )

ggplot(tox21_plot, aes(x = compound_label, y = assay, fill = value_factor)) +
  geom_tile(color = "white") +
  scale_fill_manual(
    values = c("0" = "#d1d5db", "1" = "#2563eb"),
    na.value = "#f3f4f6"
  ) +
  labs(
    title = "Tox21 outcomes (0/1) for compounds with the lowest and highest positive counts",
    x = "Drug name",
    y = "Assay",
    fill = ""
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

tox21_rate <- tox21_long |>
  group_by(assay) |>
  summarize(positive_rate = mean(value == 1, na.rm = TRUE), .groups = "drop")

ggplot(tox21_rate, aes(x = reorder(assay, positive_rate), y = positive_rate)) +
  geom_col(fill = "#4C78A8") +
  coord_flip() +
  labs(
    title = "Tox21 positive rate by assay",
    x = "Assay",
    y = "Positive rate"
  ) +
  theme_minimal()
```

## ToxCast

A compound-by-assay matrix for the ToxCast panel. Rows are assay endpoints
and columns are `HDD.Compound.ID` values. Values are typically binary (0/1) with missing values.

```{r toxcast-preview}
se <- experiments(hdd_mae)[["ToxCast"]]
assay_mat <- assays(se)[[1]]

c(rows = nrow(assay_mat), cols = ncol(assay_mat))
assay_mat[1:min(5, nrow(assay_mat)), 1:min(6, ncol(assay_mat))]
```

```{r toxcast-qc}
# Value distribution
assay_mat_num <- as.matrix(assay_mat)
table(as.vector(assay_mat_num), useNA = "ifany")

# Missingness per assay
summary(rowMeans(is.na(assay_mat_num)))
```

```{r toxcast-plots, fig.height=5, fig.width=8}
toxcast_counts <- as.data.frame(table(
  value = as.vector(assay_mat_num),
  useNA = "ifany"
))

ggplot(toxcast_counts, aes(x = value, y = Freq)) +
  geom_col(fill = "#4C78A8") +
  labs(
    title = "ToxCast value counts",
    x = "Value",
    y = "Count"
  ) +
  theme_minimal()

row_non_missing <- rowSums(!is.na(assay_mat_num))
row_positive <- rowSums(assay_mat_num == 1, na.rm = TRUE)
row_rate <- row_positive / row_non_missing
row_rate <- row_rate[!is.na(row_rate)]

ggplot(tibble::tibble(positive_rate = row_rate), aes(x = positive_rate)) +
  geom_histogram(bins = 30, fill = "#3A6EA5", alpha = 0.8) +
  labs(
    title = "ToxCast positive rate by assay",
    x = "Positive rate",
    y = "Assay count"
  ) +
  theme_minimal()

col_non_missing <- colSums(!is.na(assay_mat_num))
col_positive <- colSums(assay_mat_num == 1, na.rm = TRUE)
col_rate <- col_positive / col_non_missing
toxcast_compound <- tibble::tibble(
  compound = colnames(assay_mat_num),
  positive_rate = as.numeric(col_rate)
) |>
  left_join(compound_labels, by = c("compound" = "sample_id")) |>
  arrange(positive_rate, compound)

toxcast_top_compounds <- toxcast_compound |>
  slice_tail(n = 20) |>
  arrange(positive_rate, compound) |>
  mutate(
    display_label = ifelse(
      nchar(compound_label) > 40,
      paste0(substr(compound_label, 1, 37), "..."),
      compound_label
    ),
    display_label = factor(
      make.unique(display_label),
      levels = make.unique(display_label)
    )
  )

ggplot(
  toxcast_top_compounds,
  aes(x = display_label, y = positive_rate)
) +
  geom_col(fill = "#4C78A8") +
  labs(
    title = "ToxCast positive rate by compound (top 20)",
    x = "Drug name",
    y = "Positive rate"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1))
```

## ClinTox

A compound-by-label matrix for clinical toxicity outcomes. Rows correspond
to the ClinTox labels (e.g., FDA approval and toxicity) and columns are `HDD.Compound.ID` values.
Values are binary (0/1) with missing values for compounds not present in the source dataset.

```{r clintox-preview}
se <- experiments(hdd_mae)[["ClinTox"]]
assay_mat <- assay(se)

c(rows = nrow(assay_mat), cols = ncol(assay_mat))
assay_mat[1:min(5, nrow(assay_mat)), 1:min(6, ncol(assay_mat))]
```

```{r clintox-qc}
# Value distribution
assay_mat_num <- as.matrix(assay_mat)
table(as.vector(assay_mat_num), useNA = "ifany")

# Missingness per label
summary(rowMeans(is.na(assay_mat_num)))
```

```{r clintox-plots, fig.height=5, fig.width=8}
clintox_counts <- as.data.frame(table(
  value = as.vector(assay_mat_num),
  useNA = "ifany"
))

ggplot(clintox_counts, aes(x = value, y = Freq)) +
  geom_col(fill = "#4C78A8") +
  labs(
    title = "ClinTox value counts",
    x = "Value",
    y = "Count"
  ) +
  theme_minimal()

row_non_missing <- rowSums(!is.na(assay_mat_num))
row_positive <- rowSums(assay_mat_num == 1, na.rm = TRUE)
row_rate <- row_positive / row_non_missing

clintox_label <- tibble::tibble(
  label = rownames(assay_mat_num),
  positive_rate = as.numeric(row_rate)
)

ggplot(
  clintox_label,
  aes(x = reorder(label, positive_rate), y = positive_rate)
) +
  geom_col(fill = "#72B7B2") +
  coord_flip() +
  labs(
    title = "ClinTox positive rate by label",
    x = "Label",
    y = "Positive rate"
  ) +
  theme_minimal()
```

## SIDER

A compound-by-side-effect matrix derived from SIDER. Rows are side effect
categories and columns are `HDD.Compound.ID` values. Values are binary (0/1) with missing values when
a compound is not represented.

```{r sider-preview}
se <- experiments(hdd_mae)[["SIDER"]]
assay_mat <- assay(se)

c(rows = nrow(assay_mat), cols = ncol(assay_mat))
assay_mat[1:min(5, nrow(assay_mat)), 1:min(6, ncol(assay_mat))]
```

```{r sider-qc}
# Value distribution
assay_mat_num <- as.matrix(assay_mat)
table(as.vector(assay_mat_num), useNA = "ifany")

# Missingness per side effect
summary(rowMeans(is.na(assay_mat_num)))
```

```{r sider-plots, fig.height=6, fig.width=8}
sider_counts <- as.data.frame(table(
  value = as.vector(assay_mat_num),
  useNA = "ifany"
))

ggplot(sider_counts, aes(x = value, y = Freq)) +
  geom_col(fill = "#4C78A8") +
  labs(
    title = "SIDER value counts",
    x = "Value",
    y = "Count"
  ) +
  theme_minimal()

row_non_missing <- rowSums(!is.na(assay_mat_num))
row_positive <- rowSums(assay_mat_num == 1, na.rm = TRUE)
row_rate <- row_positive / row_non_missing

sider_effects <- tibble::tibble(
  side_effect = rownames(assay_mat_num),
  positive_rate = as.numeric(row_rate)
)

ggplot(
  sider_effects,
  aes(x = reorder(side_effect, positive_rate), y = positive_rate)
) +
  geom_col(fill = "#72B7B2") +
  coord_flip() +
  labs(
    title = "SIDER positive rate by side effect",
    x = "Side effect",
    y = "Positive rate"
  ) +
  theme_minimal()

col_non_missing <- colSums(!is.na(assay_mat_num))
col_positive <- colSums(assay_mat_num == 1, na.rm = TRUE)
col_rate <- col_positive / col_non_missing

ggplot(tibble::tibble(positive_rate = col_rate), aes(x = positive_rate)) +
  geom_histogram(bins = 12, fill = "#3A6EA5", alpha = 0.8) +
  labs(
    title = "SIDER positive rate across compounds",
    x = "Positive rate",
    y = "Compound count"
  ) +
  theme_minimal()
```

## Fingerprints (Morgan count fingerprints)

Count-based Morgan fingerprints generated from SMILES for each compound using RDKit.
Each fingerprint experiment is stored as a sparse matrix of feature counts with rows as
fingerprint bins and columns as `HDD.Compound.ID` values.

```{r fingerprints-discovery}
fp_names <- grep("^fingerprint\\.", experiment_names, value = TRUE)
fp_names

preferred_fp_names <- c(
  "fingerprint.Morgan.3.1024",
  "fingerprint.Morgan.3.2048",
  "fingerprint.Morgan.2.1024",
  "fingerprint.Morgan.2.2048"
)
selected_fp_name <- preferred_fp_names[preferred_fp_names %in% fp_names][1]
selected_fp_name
```

```{r fingerprints-preview}
for (fp_name in fp_names) {
  se <- experiments(hdd_mae)[[fp_name]]
  assay_mat <- assays(se)[[1]]

  cat("\nFingerprint:", fp_name, "\n")
  cat("Dimensions:", nrow(assay_mat), "x", ncol(assay_mat), "\n")

  # Sparsity for sparse matrices
  if (inherits(assay_mat, "sparseMatrix")) {
    sparsity <- 1 - Matrix::nnzero(assay_mat) / length(assay_mat)
    cat("Sparsity:", round(sparsity, 4), "\n")
  }

  print(assay_mat[1:min(5, nrow(assay_mat)), 1:min(6, ncol(assay_mat))])
}
```

```{r fingerprints-plots, fig.height=5, fig.width=8}
for (fp_name in fp_names) {
  se <- experiments(hdd_mae)[[fp_name]]
  assay_mat <- assays(se)[[1]]

  col_nnz <- Matrix::colSums(assay_mat > 0)
  row_nnz <- Matrix::rowSums(assay_mat > 0)

  nnz_df <- tibble::tibble(
    count = c(as.numeric(col_nnz), as.numeric(row_nnz)),
    axis = rep(c("Compound", "Feature"), c(length(col_nnz), length(row_nnz)))
  )

  p_nnz <- ggplot(nnz_df, aes(x = count, fill = axis)) +
    geom_histogram(bins = 30, alpha = 0.7, position = "identity") +
    scale_fill_brewer(palette = "Set2") +
    labs(
      title = paste("Fingerprint non-zero counts:", fp_name),
      x = "Non-zero features per entity",
      y = "Count",
      fill = ""
    ) +
    theme_minimal()

  nonzero_vals <- assay_mat@x
  val_df <- as.data.frame(table(value = nonzero_vals))

  p_vals <- ggplot(val_df, aes(x = value, y = Freq)) +
    geom_col(fill = "#3A6EA5") +
    labs(
      title = paste("Fingerprint count values:", fp_name),
      x = "Count value",
      y = "Frequency"
    ) +
    theme_minimal()

  print(p_nnz)
  print(p_vals)
}
```


## DILI and fingerprint coverage

The previous QC report included a small demonstration model that predicted DILI
severity from Morgan fingerprints. That modeling example is not a dataset QC
check and adds an optional `glmnet` dependency, so HDD v2 QC now reports the
underlying annotation and fingerprint coverage directly.

```{r dili-coverage}
col.Data <- colData(hdd_mae) |>
  as.data.frame()

dili_counts <- col.Data |>
  mutate(
    DILI.Severity = tidyr::replace_na(as.character(DILI.Severity), "Missing")
  ) |>
  count(DILI.Severity, sort = TRUE)

dili_counts

ggplot(dili_counts, aes(x = reorder(DILI.Severity, n), y = n)) +
  geom_col(fill = "#4C78A8") +
  coord_flip() +
  labs(
    title = "DILI severity annotation counts",
    x = "DILI severity",
    y = "Compound count"
  ) +
  theme_minimal()
```

```{r fingerprint-dili-coverage}
if (is.na(selected_fp_name)) {
  stop("No Morgan fingerprint experiments are available in the HDD MAE.")
}

fps <- assay(experiments(hdd_mae)[[selected_fp_name]])
cat("Using fingerprint experiment:", selected_fp_name, "\n")

fp_ids <- colnames(fps)
high_scores <- c(7, 8)
low_scores <- c(2, 3)

dili_coverage <- tibble::tibble(
  group = c(
    "High DILI severity (7/8)",
    "Low DILI severity (2/3)",
    "Any non-missing DILI severity"
  ),
  compounds = c(
    sum(col.Data$DILI.Severity %in% high_scores, na.rm = TRUE),
    sum(col.Data$DILI.Severity %in% low_scores, na.rm = TRUE),
    sum(!is.na(col.Data$DILI.Severity))
  ),
  with_selected_fingerprint = c(
    sum(
      col.Data$DILI.Severity %in%
        high_scores &
        as.character(col.Data$HDD.Compound.ID) %in% fp_ids,
      na.rm = TRUE
    ),
    sum(
      col.Data$DILI.Severity %in%
        low_scores &
        as.character(col.Data$HDD.Compound.ID) %in% fp_ids,
      na.rm = TRUE
    ),
    sum(
      !is.na(col.Data$DILI.Severity) &
        as.character(col.Data$HDD.Compound.ID) %in% fp_ids
    )
  )
)

dili_coverage
```

## Session Info

```{r session-info}
sessionInfo()
```
