---
title: "Harmonized Drug Dataset Quality Control"
author: "James Bannon, Michael Tran"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = FALSE,
  fig.width = 7,
  fig.height = 5
)
knitr::opts_knit$set(root.dir = normalizePath(".."))
options(digits = 3)
```

# Overview

This report is a quality control walkthrough for the lab-generated Harmonized
Drug Dataset stored as a `MultiAssayExperiment` (MAE). The dataset harmonizes
multiple drug-related assays and features across a shared compound universe
keyed by `HDD.Compound.ID` so downstream analyses can operate on a consistent
sample space. `Pubchem.CID` is retained when available, but it is not required
for a compound to be included.

## Data Sources

The MAE is constructed in `workflow/scripts/construct_MAE.R`. It pulls processed experiments
from `data/procdata/experiments/` and a `colData` table from `data/procdata/colData.csv`.
Key sources (from `config/pipeline.yaml` and the workflow scripts) include:

- **DeepChem datasets** (Tox21, ToxCast, SIDER, ClinTox): downloaded from DeepChem S3, then re-shaped to `HDD.Compound.ID`-by-assay matrices.
- **JUMP-CP** membership and source identifiers from the curated cpg0016 MAE.
- **OASIS** membership and source identifiers from the curated cpg0037 MAE.
- **GEOM** membership and source identifiers from the curated GEOM MAE.
- **LINCS** membership and source identifiers from the curated LINCS MAE.
- **CTRPv2** membership and master compound identifiers from the CTRPv2 PharmacoSet.
- **NCI60** membership and NSC identifiers from the NCI60-2026 PharmacoSet.
- **Bioassays**: pulled from AnnotationDB alongside compound metadata, with bioassays filtered to gold-standard assays.
- **ATC codes**: pulled from AnnotationDB when an authorization key is available.
- **LTKB**: DILI labels pulled from AnnotationDB for the LTKB Benchmark Dataset, DILIrank, and DILIst label sets.
- **LiverTox**: hepatotoxicity likelihood score and reason pulled from AnnotationDB.
- **DIRIL and DICT**: renal and cardiac toxicity labels pulled from AnnotationDB.
- **Fingerprints**: RDKit Morgan count fingerprints generated for compounds with parseable SMILES.

# Load HDD

```{r load-data}
suppressPackageStartupMessages({
  library(MultiAssayExperiment)
  library(SummarizedExperiment)
  library(Matrix)
  library(ggplot2)
  library(dplyr)
  library(tidyr)
})

hdd_mae <- readRDS("data/results/HDD_v2.3.RDS")
hdd_mae
```

```{r run-artifacts}
format_file_size <- function(bytes) {
  if (is.na(bytes)) {
    return(NA_character_)
  }

  units <- c("B", "KB", "MB", "GB")
  unit_idx <- 1
  value <- bytes
  while (value >= 1024 && unit_idx < length(units)) {
    value <- value / 1024
    unit_idx <- unit_idx + 1
  }

  sprintf("%.1f %s", value, units[[unit_idx]])
}

expected_outputs <- c(
  "data/results/HDD_v2.3.RDS",
  "data/results/HDD_v2.3_csv",
  "data/results/HDD_v2.3_csv.tar.gz"
)

output_summary <- tibble::tibble(
  path = expected_outputs,
  exists = file.exists(expected_outputs),
  size = vapply(
    expected_outputs,
    function(path) {
      if (file.exists(path) && !dir.exists(path)) {
        file.info(path)$size
      } else if (dir.exists(path)) {
        sum(
          file.info(list.files(path, recursive = TRUE, full.names = TRUE))$size,
          na.rm = TRUE
        )
      } else {
        NA_real_
      }
    },
    numeric(1)
  ),
  size_pretty = vapply(size, format_file_size, character(1)),
  modified = as.POSIXct(
    vapply(
      expected_outputs,
      function(path) {
        if (file.exists(path)) {
          file.info(path)$mtime
        } else {
          NA
        }
      },
      as.POSIXct(NA)
    ),
    origin = "1970-01-01"
  )
)

output_summary
```

```{r compound-labels}
compound_labels <- as.data.frame(colData(hdd_mae)) |>
  tibble::rownames_to_column("sample_id") |>
  transmute(
    sample_id = as.character(sample_id),
    compound_label = dplyr::coalesce(
      na_if(trimws(as.character(Molecule.Name)), ""),
      na_if(trimws(as.character(IUPAC.Name)), ""),
      na_if(trimws(as.character(Pubchem.CID)), ""),
      na_if(trimws(as.character(HDD.Compound.ID)), ""),
      sample_id
    )
  )

compound_labels$compound_label <- make.unique(compound_labels$compound_label)
```

# MAE snapshot

```{r mae-snapshot}
experiment_names <- names(experiments(hdd_mae))
experiment_names

has_experiment <- function(name) {
  name %in% experiment_names
}

assay_missing_fraction <- function(assay_obj) {
  dims <- dim(assay_obj)
  total_values <- prod(dims)

  if (inherits(assay_obj, "sparseMatrix")) {
    stored_values <- if ("x" %in% slotNames(assay_obj)) {
      slot(assay_obj, "x")
    } else {
      NULL
    }

    if (length(stored_values) == 0) {
      return(0)
    }
    return(sum(is.na(stored_values)) / total_values)
  }

  mean(is.na(assay_obj))
}

summarize_experiment <- function(se) {
  assay_obj <- assays(se)[[1]]
  data_class <- class(assay_obj)[1]
  dims <- dim(assay_obj)
  missing_fraction <- assay_missing_fraction(assay_obj)

  data.frame(
    assay_name = assayNames(se)[1],
    rows = dims[1],
    cols = dims[2],
    data_class = data_class,
    missing_fraction = missing_fraction,
    stringsAsFactors = FALSE
  )
}

experiment_summary <- do.call(
  rbind,
  lapply(experiments(hdd_mae), summarize_experiment)
)
rownames(experiment_summary) <- experiment_names
experiment_summary
```

Basic sanity checks for sampleMap completeness and the HDD schema.

```{r samplemap-checks}
# Basic sampleMap sanity checks
all_in_coldata <- lapply(experiments(hdd_mae), function(se) {
  all(colnames(se) %in% rownames(colData(hdd_mae)))
})
if (all(unlist(all_in_coldata))) {
  print("All experiments map cleanly to colData.")
} else {
  print("Some experiments contain samples missing from colData:")
  print(all_in_coldata[!unlist(all_in_coldata)])
}

# Any missing samples per experiment (only report issues)
missing_samples <- lapply(experiments(hdd_mae), function(se) {
  setdiff(colnames(se), rownames(colData(hdd_mae)))
})
missing_samples <- missing_samples[lengths(missing_samples) > 0]
if (length(missing_samples) == 0) {
  print("No missing samples detected in any experiment.")
} else {
  print("Missing samples by experiment:")
  print(missing_samples)
}

schema_checks <- tibble::tibble(
  check = c(
    "LINCS membership present",
    "JUMP-CP membership present",
    "OASIS membership present",
    "GEOM membership present",
    "CTRP membership present",
    "NCI60 membership present",
    "HDD compound ID present"
  ),
  passed = c(
    "In.LINCS" %in% colnames(colData(hdd_mae)),
    "In.JUMP.CP" %in% colnames(colData(hdd_mae)),
    "In.OASIS" %in% colnames(colData(hdd_mae)),
    "In.GEOM" %in% colnames(colData(hdd_mae)),
    "In.CTRP" %in% colnames(colData(hdd_mae)),
    "In.NCI60" %in% colnames(colData(hdd_mae)),
    "HDD.Compound.ID" %in% colnames(colData(hdd_mae))
  )
)
schema_checks
stopifnot(all(schema_checks$passed))

source_membership_counts <- tibble::tibble(
  source = c("LINCS", "JUMP-CP", "OASIS", "GEOM", "CTRPv2", "NCI60"),
  flag_column = c(
    "In.LINCS",
    "In.JUMP.CP",
    "In.OASIS",
    "In.GEOM",
    "In.CTRP",
    "In.NCI60"
  )
) |>
  mutate(
    compound_count = vapply(
      flag_column,
      function(column) {
        sum(as.logical(colData(hdd_mae)[[column]]), na.rm = TRUE)
      },
      integer(1)
    )
  )

source_membership_counts

samplemap_coverage <- tibble::tibble(
  colData_rows = nrow(colData(hdd_mae)),
  sampleMap_primary_ids = dplyr::n_distinct(sampleMap(hdd_mae)$primary),
  colData_rows_absent_from_sampleMap = sum(
    !rownames(colData(hdd_mae)) %in% sampleMap(hdd_mae)$primary
  )
)
samplemap_coverage
```

# colData overview

```{r coldata-overview}
cd <- colData(hdd_mae)
cd[1:5, 1:min(10, ncol(cd))]

# Quick missingness scan
missing_by_field <- sapply(cd, function(x) {
  mean(is.na(x))
})
missing_by_field
```

```{r coldata-basic-stats}
# Distinct counts for key identifiers
c(
  n_compounds = nrow(cd),
  unique_hdd_ids = length(unique(cd$HDD.Compound.ID)),
  unique_non_missing_cids = length(unique(na.omit(cd$Pubchem.CID)))
)
```

# Experiment walkthroughs

Below, each experiment in the HDD MAE is reviewed.

## Bioassays

A compound-by-assay matrix of PubChem bioassay outcomes. Rows are AIDs
(assay identifiers) and columns are `HDD.Compound.ID` values. Entries are categorical outcomes,
with missing values representing compounds that were not measured.

```{r bioassays-preview}
se <- experiments(hdd_mae)[["Bioassays"]]
assay_mat <- assay(se)

c(rows = nrow(assay_mat), cols = ncol(assay_mat))
assay_mat[1:min(5, nrow(assay_mat)), 1:min(6, ncol(assay_mat))]
```

```{r bioassays-qc}
# Category counts
assay_mat_chr <- as.matrix(assay_mat)
assay_values <- ifelse(
  is.na(assay_mat_chr),
  "Not Measured",
  as.character(assay_mat_chr)
)
category_counts <- sort(table(assay_values), decreasing = TRUE)
category_counts

inactive_count <- if ("Inactive" %in% names(category_counts)) {
  category_counts[["Inactive"]]
} else {
  0
}
cat("Inactive values in current Bioassays matrix:", inactive_count, "\n")

# Missingness proxy: Not Measured rate per assay
not_measured_rate <- rowMeans(is.na(assay_mat_chr))
summary(not_measured_rate)
```

```{r bioassays-plots, fig.height=6, fig.width=8}
bio_counts <- tibble::tibble(
  status = factor(
    names(category_counts),
    levels = c("Active", "Inactive", "Not Measured")
  ),
  count = as.numeric(category_counts)
)

bio_by_assay <- t(apply(assay_mat_chr, 1, function(values) {
  table(factor(
    ifelse(is.na(values), "Not Measured", as.character(values)),
    levels = c("Active", "Inactive", "Not Measured")
  ))
})) |>
  as.data.frame() |>
  tibble::rownames_to_column("assay") |>
  tidyr::pivot_longer(-assay, names_to = "status", values_to = "count")

bio_by_assay$status <- factor(
  bio_by_assay$status,
  levels = c("Active", "Inactive", "Not Measured")
)

ggplot(bio_counts, aes(x = status, y = count)) +
  geom_col(fill = "#4C78A8") +
  labs(
    title = "Bioassays overall category counts",
    x = "Outcome",
    y = "Count"
  ) +
  theme_minimal()

ggplot(bio_by_assay, aes(x = assay, y = count, fill = status)) +
  geom_col(position = "fill") +
  coord_flip() +
  labs(
    title = "Bioassays outcome composition by assay",
    x = "Assay (AID)",
    y = "Fraction",
    fill = ""
  ) +
  theme_minimal()
```

## Tox21

A compound-by-assay matrix for the Tox21 panel. Rows are Tox21 assays and
columns are `HDD.Compound.ID` values. Values are typically binary (0/1) with missing values for
unmeasured compounds.

```{r tox21-preview}
se <- experiments(hdd_mae)[["Tox21"]]
assay_mat <- assay(se)

c(rows = nrow(assay_mat), cols = ncol(assay_mat))
assay_mat[1:min(5, nrow(assay_mat)), 1:min(6, ncol(assay_mat))]
```

```{r tox21-qc}
# Value distribution
assay_mat_num <- as.matrix(assay_mat)
if ("mol_id" %in% rownames(assay_mat_num)) {
  assay_mat_num <- assay_mat_num[
    rownames(assay_mat_num) != "mol_id",
    ,
    drop = FALSE
  ]
}
table(as.vector(assay_mat_num), useNA = "ifany")

# Missingness per assay
summary(rowMeans(is.na(assay_mat_num)))
```

```{r tox21-plots, fig.height=5, fig.width=8}
tox21_long <- as.data.frame(assay_mat_num) |>
  tibble::rownames_to_column("assay") |>
  tidyr::pivot_longer(-assay, names_to = "compound", values_to = "value")
tox21_long$value_factor <- factor(tox21_long$value, levels = c(0, 1))

tox21_compound_summary <- tox21_long |>
  group_by(compound) |>
  summarize(positive_count = sum(value == 1, na.rm = TRUE), .groups = "drop") |>
  left_join(compound_labels, by = c("compound" = "sample_id")) |>
  arrange(positive_count, compound)

tox21_plot_compounds <- bind_rows(
  tox21_compound_summary |> slice_head(n = 20) |> mutate(group = "Bottom 20"),
  tox21_compound_summary |> slice_tail(n = 20) |> mutate(group = "Top 20")
) |>
  distinct(compound, .keep_all = TRUE) |>
  arrange(positive_count, compound)

tox21_plot <- tox21_long |>
  filter(compound %in% tox21_plot_compounds$compound) |>
  left_join(
    tox21_plot_compounds |>
      select(compound, compound_label),
    by = "compound"
  ) |>
  mutate(
    compound_label = factor(
      compound_label,
      levels = tox21_plot_compounds$compound_label
    )
  )

ggplot(tox21_plot, aes(x = compound_label, y = assay, fill = value_factor)) +
  geom_tile(color = "white") +
  scale_fill_manual(
    values = c("0" = "#d1d5db", "1" = "#2563eb"),
    na.value = "#f3f4f6"
  ) +
  labs(
    title = "Tox21 outcomes (0/1) for compounds with the lowest and highest positive counts",
    x = "Drug name",
    y = "Assay",
    fill = ""
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

tox21_rate <- tox21_long |>
  group_by(assay) |>
  summarize(positive_rate = mean(value == 1, na.rm = TRUE), .groups = "drop")

ggplot(tox21_rate, aes(x = reorder(assay, positive_rate), y = positive_rate)) +
  geom_col(fill = "#4C78A8") +
  coord_flip() +
  labs(
    title = "Tox21 positive rate by assay",
    x = "Assay",
    y = "Positive rate"
  ) +
  theme_minimal()
```

## ToxCast

A compound-by-assay matrix for the ToxCast panel. Rows are assay endpoints
and columns are `HDD.Compound.ID` values. Values are typically binary (0/1) with missing values.

```{r toxcast-preview}
se <- experiments(hdd_mae)[["ToxCast"]]
assay_mat <- assays(se)[[1]]

c(rows = nrow(assay_mat), cols = ncol(assay_mat))
assay_mat[1:min(5, nrow(assay_mat)), 1:min(6, ncol(assay_mat))]
```

```{r toxcast-qc}
# Value distribution
assay_mat_num <- as.matrix(assay_mat)
table(as.vector(assay_mat_num), useNA = "ifany")

# Missingness per assay
summary(rowMeans(is.na(assay_mat_num)))
```

```{r toxcast-plots, fig.height=5, fig.width=8}
toxcast_counts <- as.data.frame(table(
  value = as.vector(assay_mat_num),
  useNA = "ifany"
))

ggplot(toxcast_counts, aes(x = value, y = Freq)) +
  geom_col(fill = "#4C78A8") +
  labs(
    title = "ToxCast value counts",
    x = "Value",
    y = "Count"
  ) +
  theme_minimal()

row_non_missing <- rowSums(!is.na(assay_mat_num))
row_positive <- rowSums(assay_mat_num == 1, na.rm = TRUE)
row_rate <- row_positive / row_non_missing
row_rate <- row_rate[!is.na(row_rate)]

ggplot(tibble::tibble(positive_rate = row_rate), aes(x = positive_rate)) +
  geom_histogram(bins = 30, fill = "#3A6EA5", alpha = 0.8) +
  labs(
    title = "ToxCast positive rate by assay",
    x = "Positive rate",
    y = "Assay count"
  ) +
  theme_minimal()

col_non_missing <- colSums(!is.na(assay_mat_num))
col_positive <- colSums(assay_mat_num == 1, na.rm = TRUE)
col_rate <- col_positive / col_non_missing
toxcast_compound <- tibble::tibble(
  compound = colnames(assay_mat_num),
  positive_rate = as.numeric(col_rate)
) |>
  left_join(compound_labels, by = c("compound" = "sample_id")) |>
  arrange(positive_rate, compound)

toxcast_top_compounds <- toxcast_compound |>
  slice_tail(n = 20) |>
  arrange(positive_rate, compound) |>
  mutate(
    display_label = ifelse(
      nchar(compound_label) > 40,
      paste0(substr(compound_label, 1, 37), "..."),
      compound_label
    ),
    display_label = factor(
      make.unique(display_label),
      levels = make.unique(display_label)
    )
  )

ggplot(
  toxcast_top_compounds,
  aes(x = display_label, y = positive_rate)
) +
  geom_col(fill = "#4C78A8") +
  labs(
    title = "ToxCast positive rate by compound (top 20)",
    x = "Drug name",
    y = "Positive rate"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1))
```

## ClinTox

A compound-by-label matrix for clinical toxicity outcomes. Rows correspond
to the ClinTox labels (e.g., FDA approval and toxicity) and columns are `HDD.Compound.ID` values.
Values are binary (0/1) with missing values for compounds not present in the source dataset.

```{r clintox-preview}
se <- experiments(hdd_mae)[["ClinTox"]]
assay_mat <- assay(se)

c(rows = nrow(assay_mat), cols = ncol(assay_mat))
assay_mat[1:min(5, nrow(assay_mat)), 1:min(6, ncol(assay_mat))]
```

```{r clintox-qc}
# Value distribution
assay_mat_num <- as.matrix(assay_mat)
table(as.vector(assay_mat_num), useNA = "ifany")

# Missingness per label
summary(rowMeans(is.na(assay_mat_num)))
```

```{r clintox-plots, fig.height=5, fig.width=8}
clintox_counts <- as.data.frame(table(
  value = as.vector(assay_mat_num),
  useNA = "ifany"
))

ggplot(clintox_counts, aes(x = value, y = Freq)) +
  geom_col(fill = "#4C78A8") +
  labs(
    title = "ClinTox value counts",
    x = "Value",
    y = "Count"
  ) +
  theme_minimal()

row_non_missing <- rowSums(!is.na(assay_mat_num))
row_positive <- rowSums(assay_mat_num == 1, na.rm = TRUE)
row_rate <- row_positive / row_non_missing

clintox_label <- tibble::tibble(
  label = rownames(assay_mat_num),
  positive_rate = as.numeric(row_rate)
)

ggplot(
  clintox_label,
  aes(x = reorder(label, positive_rate), y = positive_rate)
) +
  geom_col(fill = "#72B7B2") +
  coord_flip() +
  labs(
    title = "ClinTox positive rate by label",
    x = "Label",
    y = "Positive rate"
  ) +
  theme_minimal()
```

## SIDER

A compound-by-side-effect matrix derived from SIDER. Rows are side effect
categories and columns are `HDD.Compound.ID` values. Values are binary (0/1) with missing values when
a compound is not represented.

```{r sider-preview}
se <- experiments(hdd_mae)[["SIDER"]]
assay_mat <- assay(se)

c(rows = nrow(assay_mat), cols = ncol(assay_mat))
assay_mat[1:min(5, nrow(assay_mat)), 1:min(6, ncol(assay_mat))]
```

```{r sider-qc}
# Value distribution
assay_mat_num <- as.matrix(assay_mat)
table(as.vector(assay_mat_num), useNA = "ifany")

# Missingness per side effect
summary(rowMeans(is.na(assay_mat_num)))
```

```{r sider-plots, fig.height=6, fig.width=8}
sider_counts <- as.data.frame(table(
  value = as.vector(assay_mat_num),
  useNA = "ifany"
))

ggplot(sider_counts, aes(x = value, y = Freq)) +
  geom_col(fill = "#4C78A8") +
  labs(
    title = "SIDER value counts",
    x = "Value",
    y = "Count"
  ) +
  theme_minimal()

row_non_missing <- rowSums(!is.na(assay_mat_num))
row_positive <- rowSums(assay_mat_num == 1, na.rm = TRUE)
row_rate <- row_positive / row_non_missing

sider_effects <- tibble::tibble(
  side_effect = rownames(assay_mat_num),
  positive_rate = as.numeric(row_rate)
)

ggplot(
  sider_effects,
  aes(x = reorder(side_effect, positive_rate), y = positive_rate)
) +
  geom_col(fill = "#72B7B2") +
  coord_flip() +
  labs(
    title = "SIDER positive rate by side effect",
    x = "Side effect",
    y = "Positive rate"
  ) +
  theme_minimal()

col_non_missing <- colSums(!is.na(assay_mat_num))
col_positive <- colSums(assay_mat_num == 1, na.rm = TRUE)
col_rate <- col_positive / col_non_missing

ggplot(tibble::tibble(positive_rate = col_rate), aes(x = positive_rate)) +
  geom_histogram(bins = 12, fill = "#3A6EA5", alpha = 0.8) +
  labs(
    title = "SIDER positive rate across compounds",
    x = "Positive rate",
    y = "Compound count"
  ) +
  theme_minimal()
```

## Fingerprints (Morgan count fingerprints)

Count-based Morgan fingerprints generated from parseable SMILES using RDKit.
Each fingerprint experiment is stored as a sparse matrix of feature counts with rows as
fingerprint bins and columns as the `HDD.Compound.ID` values that produced a fingerprint.
Compounds with missing or invalid SMILES remain in `colData` but are not columns in the
fingerprint experiments.

```{r fingerprints-discovery}
fp_names <- grep("^fingerprint\\.", experiment_names, value = TRUE)
fp_names

preferred_fp_names <- c(
  "fingerprint.Morgan.3.1024",
  "fingerprint.Morgan.3.2048",
  "fingerprint.Morgan.2.1024",
  "fingerprint.Morgan.2.2048"
)
selected_fp_name <- preferred_fp_names[preferred_fp_names %in% fp_names][1]
selected_fp_name
```

```{r fingerprints-preview}
fingerprint_coverage <- tibble::tibble(
  total_compounds = nrow(cd),
  compounds_with_smiles = sum(!is.na(cd$SMILES)),
  selected_fingerprint_columns = if (is.na(selected_fp_name)) {
    NA_integer_
  } else {
    ncol(assay(experiments(hdd_mae)[[selected_fp_name]]))
  },
  not_fingerprinted = total_compounds - selected_fingerprint_columns
)

fingerprint_coverage

for (fp_name in fp_names) {
  se <- experiments(hdd_mae)[[fp_name]]
  assay_mat <- assays(se)[[1]]

  cat("\nFingerprint:", fp_name, "\n")
  cat("Dimensions:", nrow(assay_mat), "x", ncol(assay_mat), "\n")

  # Sparsity for sparse matrices
  if (inherits(assay_mat, "sparseMatrix")) {
    sparsity <- 1 - Matrix::nnzero(assay_mat) / length(assay_mat)
    cat("Sparsity:", round(sparsity, 4), "\n")
  }

  print(assay_mat[1:min(5, nrow(assay_mat)), 1:min(6, ncol(assay_mat))])
}
```

```{r fingerprints-plots, fig.height=5, fig.width=8}
for (fp_name in fp_names) {
  se <- experiments(hdd_mae)[[fp_name]]
  assay_mat <- assays(se)[[1]]

  col_nnz <- Matrix::colSums(assay_mat > 0)
  row_nnz <- Matrix::rowSums(assay_mat > 0)

  nnz_df <- tibble::tibble(
    count = c(as.numeric(col_nnz), as.numeric(row_nnz)),
    axis = rep(c("Compound", "Feature"), c(length(col_nnz), length(row_nnz)))
  )

  p_nnz <- ggplot(nnz_df, aes(x = count, fill = axis)) +
    geom_histogram(bins = 30, alpha = 0.7, position = "identity") +
    scale_fill_brewer(palette = "Set2") +
    labs(
      title = paste("Fingerprint non-zero counts:", fp_name),
      x = "Non-zero features per entity",
      y = "Count",
      fill = ""
    ) +
    theme_minimal()

  nonzero_vals <- assay_mat@x
  val_df <- as.data.frame(table(value = nonzero_vals))

  p_vals <- ggplot(val_df, aes(x = value, y = Freq)) +
    geom_col(fill = "#3A6EA5") +
    labs(
      title = paste("Fingerprint count values:", fp_name),
      x = "Count value",
      y = "Frequency"
    ) +
    theme_minimal()

  print(p_nnz)
  print(p_vals)
}
```


## ATC code coverage

AnnotationDB ATC data requires an authorization key. If the key is not set,
the pipeline still builds the HDD with other AnnotationDB fields and leaves
`ATC.Code` missing.

```{r atc-code-coverage, fig.height=3.5, fig.width=7}
stopifnot("ATC.Code" %in% colnames(cd))

atc_values <- trimws(as.character(cd$ATC.Code))
atc_coverage <- tibble::tibble(
  status = c("ATC code present", "ATC code missing"),
  compounds = c(
    sum(!is.na(atc_values) & atc_values != ""),
    sum(is.na(atc_values) | atc_values == "")
  )
)

ggplot(atc_coverage, aes(x = status, y = compounds, fill = status)) +
  geom_col(width = 0.65, show.legend = FALSE) +
  geom_text(aes(label = scales::comma(compounds)), vjust = -0.4) +
  scale_y_continuous(
    labels = scales::comma,
    expand = expansion(mult = c(0, 0.12))
  ) +
  scale_fill_manual(values = c(
    "ATC code present" = "#26734D",
    "ATC code missing" = "#BDBDBD"
  )) +
  labs(
    title = "ATC code coverage in HDD compound metadata",
    x = NULL,
    y = "Compounds"
  ) +
  theme_minimal(base_size = 11)
```

## DILI, DIRIL, and DICT coverage

This section summarizes the toxicity metadata currently available in `colData`
for liver, renal, and cardiac toxicity labels.

```{r source-toxicity-coverage}
col.Data <- colData(hdd_mae) |>
  as.data.frame()

source_toxicity_columns <- tibble::tribble(
  ~source, ~label_set, ~field, ~column,
  "LTKB", "LTKB Benchmark Dataset", "DILI severity grade",
  "LTKB.DILI.Severity.Grade",
  "LTKB", "LTKB Benchmark Dataset", "DILI annotation", "LTKB.DILI.Annotation",
  "LTKB", "DILIrank", "DILI severity grade", "DILIrank.DILI.Severity.Grade",
  "LTKB", "DILIrank", "DILI annotation", "DILIrank.DILI.Annotation",
  "LTKB", "DILIst", "DILI annotation", "DILIst.DILI.Annotation",
  "LiverTox", "LiverTox", "hepatotoxicity likelihood score",
  "Livertox.Hepatotoxicity.Likelihood.Score",
  "LiverTox", "LiverTox", "hepatotoxicity likelihood reason",
  "Livertox.Hepatotoxicity.Likelihood.Reason"
)

diril_dict_columns <- tibble::tribble(
  ~source, ~field, ~column,
  "DIRIL", "Gong label", "DIRIL.Label.Gong",
  "DIRIL", "Shi label", "DIRIL.Label.Shi",
  "DIRIL", "final toxicity label", "DIRIL.Toxicity",
  "DICT", "cardiotoxicity class", "DICT.Cardiotoxicity",
  "DICT", "label section", "DICT.Label.Section",
  "DICT", "concern", "DICT.Concern",
  "DICT", "keywords", "DICT.Keywords"
)

count_non_missing <- function(column) {
  values <- as.character(col.Data[[column]])
  sum(!is.na(values) & trimws(values) != "")
}

count_unique_non_missing <- function(column) {
  values <- as.character(col.Data[[column]])
  non_missing_values <- values[!is.na(values) & trimws(values) != ""]
  length(unique(non_missing_values))
}

source_toxicity_coverage <- source_toxicity_columns |>
  mutate(
    non_missing = vapply(column, count_non_missing, integer(1)),
    unique_non_missing = vapply(column, count_unique_non_missing, integer(1))
  )

source_toxicity_coverage |>
  select(source, label_set, field, non_missing, unique_non_missing)

ggplot(
  source_toxicity_coverage |>
    mutate(label = paste(label_set, field, sep = ": ")),
  aes(x = reorder(label, non_missing), y = non_missing, fill = source)
) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Non-missing LTKB and LiverTox metadata",
    x = "Label set and field",
    y = "Non-missing compound count",
    fill = "Source"
  ) +
  theme_minimal()
```

```{r toxicity-label-counts}
toxicity_label_count_columns <- bind_rows(
  source_toxicity_columns |>
    filter(field != "hepatotoxicity likelihood reason"),
  diril_dict_columns |>
    filter(field != "keywords")
)

toxicity_label_counts <- lapply(
  seq_len(nrow(toxicity_label_count_columns)),
  function(idx) {
    spec <- toxicity_label_count_columns[idx, ]
    values <- trimws(as.character(col.Data[[spec$column]]))

    tibble::tibble(
      source = spec$source,
      label_set = spec$label_set,
      field = spec$field,
      label = values
    ) |>
      filter(!is.na(label), label != "") |>
      count(source, label_set, field, label, sort = TRUE)
  }
) |>
  dplyr::bind_rows()

toxicity_label_counts
```

```{r dili-coverage}
source_dili_severity_columns <- source_toxicity_columns |>
  filter(field == "DILI severity grade") |>
  pull(column)

dili_counts <- lapply(source_dili_severity_columns, function(column) {
  label_set_name <- source_toxicity_columns$label_set[
    match(column, source_toxicity_columns$column)
  ]

  tibble::tibble(
    label_set = label_set_name,
    column = column,
    severity_grade = trimws(as.character(col.Data[[column]]))
  ) |>
    filter(!is.na(severity_grade), severity_grade != "") |>
    count(label_set, column, severity_grade, sort = TRUE)
}) |>
  dplyr::bind_rows() |>
  mutate(
    severity_grade = factor(
      severity_grade,
      levels = sort(unique(as.numeric(severity_grade)))
    )
  )

dili_counts

ggplot(dili_counts, aes(x = severity_grade, y = n, fill = label_set)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.72) +
  labs(
    title = "DILI severity annotation counts",
    x = "DILI severity grade",
    y = "Compound count",
    fill = "Label set"
  ) +
  theme_minimal()
```

```{r dili-severity-consensus}
dili_severity_consensus <- col.Data |>
  transmute(
    LTKB = suppressWarnings(
      as.integer(trimws(as.character(LTKB.DILI.Severity.Grade)))
    ),
    DILIrank = suppressWarnings(
      as.integer(trimws(as.character(DILIrank.DILI.Severity.Grade)))
    )
  ) |>
  filter(!is.na(LTKB), !is.na(DILIrank)) |>
  count(LTKB, DILIrank, name = "n") |>
  tidyr::complete(
    LTKB = seq(min(LTKB), max(LTKB)),
    DILIrank = seq(min(DILIrank), max(DILIrank)),
    fill = list(n = 0L)
  )

ggplot(dili_severity_consensus, aes(x = LTKB, y = DILIrank, fill = n)) +
  geom_tile(color = "white", linewidth = 0.4) +
  geom_text(aes(label = ifelse(n > 0, n, "")), size = 3) +
  scale_x_continuous(breaks = sort(unique(dili_severity_consensus$LTKB))) +
  scale_y_continuous(breaks = sort(unique(dili_severity_consensus$DILIrank))) +
  scale_fill_gradient(low = "#EDF5FA", high = "#2F5D7C") +
  labs(
    title = "LTKB Benchmark Dataset and DILIrank severity grade agreement",
    x = "LTKB Benchmark Dataset severity grade",
    y = "DILIrank severity grade",
    fill = "Compound count"
  ) +
  theme_minimal()
```

```{r diril-dict-coverage}
diril_dict_coverage <- diril_dict_columns |>
  mutate(
    non_missing = vapply(column, count_non_missing, integer(1)),
    unique_non_missing = vapply(column, count_unique_non_missing, integer(1))
  )

diril_dict_coverage
```

```{r chembl-moa-toxicity-labels, fig.width=11, fig.height=14}
moa_toxicity_label_columns <- tibble::tribble(
  ~source, ~label_set, ~field, ~column,
  "LTKB", "LTKB Benchmark Dataset", "DILI annotation", "LTKB.DILI.Annotation",
  "LTKB", "DILIrank", "DILI annotation", "DILIrank.DILI.Annotation",
  "LTKB", "DILIst", "DILI annotation", "DILIst.DILI.Annotation",
  "LiverTox", "LiverTox", "hepatotoxicity likelihood score",
  "Livertox.Hepatotoxicity.Likelihood.Score",
  "DIRIL", "DIRIL", "final toxicity label", "DIRIL.Toxicity",
  "DICT", "DICT", "concern", "DICT.Concern"
)

moa_toxicity_labels <- lapply(
  seq_len(nrow(moa_toxicity_label_columns)),
  function(idx) {
    spec <- moa_toxicity_label_columns[idx, ]
    moa <- trimws(as.character(col.Data$Mechanism.of.Action))
    label <- trimws(as.character(col.Data[[spec$column]]))

    tibble::tibble(
      source = spec$source,
      label_set = spec$label_set,
      field = spec$field,
      moa = moa,
      toxicity_label = label
    ) |>
      filter(
        !is.na(moa),
        moa != "",
        !is.na(toxicity_label),
        toxicity_label != ""
      )
  }
) |>
  dplyr::bind_rows()

wrap_plot_label <- function(value, width = 34) {
  vapply(
    value,
    function(item) {
      item <- trimws(as.character(item))
      if (is.na(item) || item == "") {
        return(NA_character_)
      }
      paste(strwrap(item, width = width), collapse = "\n")
    },
    character(1),
    USE.NAMES = FALSE
  )
}

moa_toxicity_counts <- moa_toxicity_labels |>
  count(source, label_set, field, moa, toxicity_label, sort = TRUE)

n_top_moa_toxicity_pairs <- 6

moa_toxicity_plot_data <- moa_toxicity_counts |>
  arrange(source, label_set, desc(n), moa, toxicity_label) |>
  group_by(source, label_set) |>
  slice_head(n = n_top_moa_toxicity_pairs) |>
  mutate(
    pair_rank = row_number(),
    pair_label = wrap_plot_label(moa, 42),
    toxicity_label_plot = wrap_plot_label(toxicity_label, 28),
    panel = paste(source, label_set, sep = ": "),
    pair_key = paste(panel, pair_rank, pair_label, sep = "|||")
  ) |>
  ungroup() |>
  arrange(source, label_set, desc(pair_rank)) |>
  mutate(
    source = factor(source, levels = unique(moa_toxicity_label_columns$source)),
    panel = factor(panel, levels = unique(panel)),
    pair_key = factor(pair_key, levels = unique(pair_key)),
    toxicity_label_plot = factor(
      toxicity_label_plot,
      levels = sort(unique(toxicity_label_plot))
    )
  )

toxicity_label_palette <- setNames(
  grDevices::hcl.colors(
    n = length(levels(moa_toxicity_plot_data$toxicity_label_plot)),
    palette = "Dark 3"
  ),
  levels(moa_toxicity_plot_data$toxicity_label_plot)
)

ggplot(
  moa_toxicity_plot_data,
  aes(x = n, y = pair_key, fill = toxicity_label_plot)
) +
  geom_col(width = 0.72) +
  geom_text(aes(label = n), hjust = -0.15, size = 2.8) +
  facet_wrap(~panel, scales = "free_y", ncol = 1) +
  scale_fill_manual(values = toxicity_label_palette) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.1))) +
  scale_y_discrete(labels = function(value) sub("^.*\\|\\|\\|", "", value)) +
  guides(fill = guide_legend(ncol = 2, byrow = TRUE)) +
  labs(
    title = "Top ChEMBL mechanism-of-action and toxicity-label combinations",
    x = "Compound count",
    y = NULL,
    fill = "Toxicity label"
  ) +
  coord_cartesian(clip = "off") +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 8, lineheight = 0.95),
    legend.position = "bottom",
    legend.text = element_text(size = 8, lineheight = 0.95),
    legend.title = element_text(size = 9),
    panel.grid.major.y = element_blank(),
    plot.margin = margin(5.5, 28, 5.5, 5.5),
    strip.text = element_text(face = "bold", hjust = 0)
  )
```

```{r diril-consensus}
normalize_diril_label <- function(value) {
  value <- trimws(as.character(value))
  value[value == ""] <- NA_character_
  dplyr::case_when(
    is.na(value) ~ NA_character_,
    tolower(value) == "nephrotoxic" ~ "Nephrotoxic",
    tolower(value) == "non-nephrotoxic" ~ "Non-Nephrotoxic",
    TRUE ~ value
  )
}

diril_consensus <- col.Data |>
  transmute(
    Gong = normalize_diril_label(DIRIL.Label.Gong),
    Shi = normalize_diril_label(DIRIL.Label.Shi),
    Final = normalize_diril_label(DIRIL.Toxicity)
  ) |>
  filter(!is.na(Gong), !is.na(Shi), !is.na(Final)) |>
  count(Gong, Shi, Final, sort = TRUE)

diril_consensus

ggplot(diril_consensus, aes(x = Gong, y = Shi, fill = n)) +
  geom_tile(color = "white", linewidth = 0.4) +
  geom_text(aes(label = n), size = 3) +
  facet_wrap(~Final) +
  scale_fill_gradient(low = "#D7E8F5", high = "#2F5D7C") +
  labs(
    title = "DIRIL Gong and Shi labels by final toxicity label",
    x = "Gong label",
    y = "Shi label",
    fill = "Compound count"
  ) +
  theme_minimal()
```

## Session Info

```{r session-info}
sessionInfo()
```
